---
title: "Data Mining - Chia Sheng Tu - HW1"
output: md_document
date: "2023-01-30"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(rsample)
library(caret)
library(modelr)
library(parallel)
library(foreach)
library(dplyr)
```
## 1) Data visualization: flights at ABIA
****Question: What is the best Month to fly in and fly out to Austin to minimize delays?**
```{r problem 1, message=FALSE, echo=FALSE, warning=FALSE}
ABIA = read.csv("Data\\ABIA.csv")
# To find the minimize delay time 
# Set average delay time per mile = (ArrDelay + DepDelay + WeatherDelay)/Distance

# Group the data by month (1 - 12)
month = ABIA %>% group_by(Month)

# Replace N/A values with 0s 
adj_Month <- replace(month, is.na(month), 0)

# Set average delay time per mile = (ArrDelay + DepDelay + WeatherDelay)/Distance
# ArrDelay arrival delay, in minutes
# DepDelay departure delay, in minutes
# WeatherDelay in minutes
rank <- adj_Month %>%
  summarize(avg_Delay = mean((ArrDelay + DepDelay + WeatherDelay)/Distance)) %>%
  round(3)

# check which month has the lowest delay time per mile in average
rank %>% head(desc(5))
# September would be the best month to fly, with minimal delay time 0.09 minutes per mile in average.

# Plot for average delay time per mile bt month 
ggplot(rank) + 
  geom_col(aes(x=factor(Month), y=avg_Delay))  + 
  labs(x="Month",
       
       y="Minutes",
       
       title="Average Delay Time per mile by Month") 

```

```
**Figure 1:** #Plot for average delay time per mile bt month

Figure 1 refers to the best months to fly in and fly out to  Austin to minimize delays are September and November. September has the lowest delay time per mile (0.009). On the Other hand, December would be the worst time to fly, since the average delay time per mile is the highst among 12 months (0.077). The results seems reasonible. The demnd of flights in September might decreas, as September is the second month of the semester for most students. However, many people, not only students, are flying in/out to Austin due to Christmas and new year hoildays, the demand of flights increased.
```

## 2) Wrangling the Olympics

#### A) What is 95th percentile of heights for female competitiors across Athletics events?
```{r problem 2A, message=FALSE, echo=FALSE}
olympics_top20 = read.csv("Data\\olympics_top20.csv")
# Filter out data that match gender female and sport Athletics
olympics_top20_filter <- olympics_top20 %>%
  filter(sex == "F" & sport == "Athletics") %>% 
  group_by(event) 
# Get 95th percentile of heights
Output = quantile(olympics_top20_filter$height, 0.95)
Numquantile="95%"
df1<- data.frame(Numquantile,Output, row.names = "output")
knitr::kable(head(df1, 1), col.names = c("quantile", "height"))
```

95th percentile of heights for female competitors across Athletcis events is 183cm.

#### B) Which single women's event had the greatest variability in competitor's heights across the entire history of the Olympics, as measured by the standard deviation?
```{r problem 2B, message=FALSE, echo=FALSE}
# Filter out data that match gender female, and pick variables event and height
# Check the standard deviation across all events
gender2 <- olympics_top20 %>%
  filter(sex == "F") %>%
  select(event, height) %>%
  group_by(event) %>%
  summarize(sd_height = sd(height)) %>% 
  arrange(desc(sd_height)) %>% 
  head()
gender2  

```

Event Rowing Woman's Coxed Fours has the highest standard deviation with 10.9

#### C) How has the average age of Olympic swimmers changed over time? Does this trend look different depending on the gender?
```{r problem 2C, message=FALSE, echo=FALSE}
# Filter out data that match sport swimming, and calculate the mean of age across all swimming sports 
age <- olympics_top20 %>%
  filter(sport == "Swimming") %>%
  group_by(year) %>% 
  summarize(mean_age = mean(age))

# Plot for Average Age of Olympic Swimmers Changed Over Time
ggplot(age) + 
  geom_line(aes(x = year, y = mean_age)) +
  labs(x="Year", y = "Average Age",
       title="Average Age of Olympic Swimmers Changed Over Time")

# Group the data by gender
agetest <- olympics_top20 %>%
  filter(sport == "Swimming") %>%
  group_by(sex, year) %>% 
  summarize(mean_age = mean(age))

# Plot for Average Age of Olympic Swimmers Changed Over Time based on gender
ggplot(agetest) + 
  geom_line(aes(x = year, y = mean_age, color = sex)) + 
  labs(x="Year", y = "Average Age",
       title="Average Age of Male/Female 
Olympic Swimmers Changed Over Time")

```
**Figure 3&4:** Line graph showing the change in average age of Olympic swimmers between 1600 - 2016
How has the average age of Olympic swimmers changed over time? 
Compared to early 1900's, the average age of Olympic swimmers has became bigger in 2000's. The average age in the early 1900's has the lowest value, and kept increasing until around 1920. After 1920, the the average age of Olympic swimmers stared to decreased until 1975.

Does the trend look different for male swimmers relative to female swimmers? 
Average age of female swimmers is lower than that of male swimmers throughout history.  Also, number of male swimmer was higher than female swimmers from 1900 to 1925, since the pattern in Figure 3 in primary is affected by male swimmers, as Figure 4 has showed.


## 3) K-nearest neighbors: cars

### Car trim level 350
```{r problem 3_1.1, message=FALSE, echo=FALSE}
sclass = read.csv("Data\\sclass.csv")

class_350 <- sclass %>% 
  filter(trim == "350")

sclass350_split = initial_split(class_350, prop=0.8)
sclass350_train = training(sclass350_split)
sclass350_test = testing(sclass350_split)

k_grid = 2:150
rmse_350_out = foreach(i=2:150, .combine = 'c') %do% {
  knn_ = knnreg(price ~ mileage, data=sclass350_train, k = i)
  rmse(knn_, sclass350_test)
}

# Data frame for plotting
rmse_350_new <- rmse_350_out %>%
  as.data.frame()

ggplot(rmse_350_new, aes(x = k_grid, y = rmse_350_out)) +
  geom_line(color = "red") +
  labs(x="k", y="RMSE",
       title = "Car's trim level 350: RMSE vs value of K >= 2") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```
**Figure 5:** Plot showing RMSE for each value of K (from 2 to 150)

Here, we can see which K value has the lowest RMSE, and this is the optimal value of K

**Optimal value of K**

```{r problem 3_1.2, message=FALSE, echo=FALSE}

# optimal value of K = 5
k_grid[which.min(rmse_350_out)]

```

**RMSE from the Optimal value of K**

```{r problem 3_1.3, message=FALSE, echo=FALSE}

# RMSE of optimal K we found
knn_350_RMSE = knnreg(price ~ mileage, data=sclass350_train, k=k_grid[which.min(rmse_350_out)])
rmse(knn_350_RMSE,sclass350_test)

```


```{r problem 3_1.4, message=FALSE, echo=FALSE}
sclass350_test = sclass350_test %>%
  mutate(price_pred = predict(knn_350_RMSE, sclass350_test))
ptest = ggplot(data = sclass350_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.5) +
  ggtitle("Predictions of price given mileage")
# Add the predictions
ptest + 
  geom_line(aes(x = mileage, y = price_pred), color='red', size=1.5) +
  theme_minimal() +
  theme(plot.title = element_text(face="bold"))

```

**Figure 2:** Plot of the fitted model, predicted price and real price


### Car trim level 65 AMG

```{r problem 3_2.1, message=FALSE, echo=FALSE}
AMG_65 <- sclass %>% 
  filter(trim == "65 AMG")
# Split data to train&testing sets 
sclass65_split = initial_split(AMG_65, prop=0.8)
sclass65_train = training(sclass65_split)
sclass65_test = testing(sclass65_split)

k_grid = 2:150
rmse_65_out = foreach(i=2:150, .combine = 'c') %do% {
  knn_65 = knnreg(price ~ mileage, data=sclass65_train, k = i)
  rmse(knn_65, sclass65_test)
}

# data frame for plotting
rmse_65_new <- rmse_65_out %>%
  as.data.frame()

ggplot(rmse_65_new, aes(x = k_grid, y = rmse_65_out)) +
  geom_line(color = "red") +
  labs(x="k", y="RMSE",
       title = "Car's trim level 65 AMG: RMSE vs value of K >= 2") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))
```


**Figure 7:** Plot showing RMSE for each value of K (from 2 to 150)

Here, we can see which K value has the lowest RMSE, and this is the optimal value of K

**Optimal value of K**

```{r problem 3_2.2, message=FALSE, echo=FALSE}

# optimal value of K = 17
k_grid[which.min(rmse_65_out)]

```

**RMSE from the Optimal value of K**

```{r problem 3_2.3, message=FALSE, echo=FALSE}

# RMSE of optimal K we found
knn_65_RMSE = knnreg(price ~ mileage, data=sclass65_train, k=k_grid[which.min(rmse_65_out)])
rmse(knn_65_RMSE,sclass65_test)

```

```{r problem 3_2.4, message=FALSE, echo=FALSE}

sclass65_test = sclass65_test %>%
  mutate(price_pred = predict(knn_65_RMSE, sclass65_test))
ptest = ggplot(data = sclass65_test) + 
  geom_point(mapping = aes(x = mileage, y = price), alpha=0.5) +
  ggtitle("Predictions of price given mileage")
# now add the predictions
ptest + 
  geom_line(aes(x = mileage, y = price_pred), color='red', size=1.5) +
  theme_minimal() +
  theme(plot.title = element_text(face="bold"))

```
**Figure 8:** Plot of the fitted model, predicted price and real price


**Result: Trim 350 yields a larger optimal value of K**
```{r problem 3_3, message=FALSE, echo=FALSE}
Trim_350_row=length(which(sclass$trim=="350"))
Trim_65_row=length(which(sclass$trim=="65 AMG"))
df<- data.frame(Trim_350_row,Trim_65_row, row.names = "Number of rows")
knitr::kable(head(df, 1), col.names = c("Trim 350", "Trim 65 AMG"))
```

Trim 350 has larger optimal K value as it contains more data points than trim 65 AMG. 
More data points means lower variance, less chance of memorizing random noise. 